\abbreviations

% You can put here what you like, but here's an example
%Note the use of starred section commands here to produce proper division
%headers without bad '0.1' numbers or entries into the Table of Contents.
%Using the {\verb \begin{symbollist} } environment ensures that entries are
%properly spaced.

%\section*{Symbols}
%\begin{symbollist}
%	\item[$\mathbb{X}$] A blackboard bold $X$.  Neat.
	% Optional item argument makes the symbol/abbr
%	\item[$\mathcal{X}$] A caligraphic $X$.  Neat.
%	\item[$\mathfrak{X}$] A fraktur $X$.  Neat.
%	\item[$\mathbf{X}$] A boldface $X$.
%	\item[$\mathsf{X}$] A sans-serif $X$. Bad notation.
%	\item[$\mathrm{X}$] A roman $X$.
%\end{symbollist}

\section*{Abbreviations}

\begin{symbollist}
	\item[VM] Vascular Malformation, a blood vessel that is malformed, though functional.
	
	\item[CCM] Cerebral Cavernous Malformation, a common vascular malformation that occurs in the brain and spinal cord. As is standard in the CCM research community, the abbreviation `CCM' is used to refer to the familial disorder, the vascular malformation itself, and occasionally the `CCM Genes' (referring to \italicize{KRIT1}, \italicize{CCM2}, and \italicize{PDCD10}) and the `CCM Complex' (the signaling complex formed by the proteins: KRIT1, CCM2, and PDCD10). 
	
	\item[HHT] Hereditary Hemorrhagic Telangiectasia, an inherited vascular malformation disorder characterized by telangiectasia and arteriovenous malformations. 
	
	\item[AVM] Arteriovenous Malformation, a vascular malformation consisting of a direct shunt between an artery and vein, bypassing the capillary bed. 
	
	\item[DVA] Developmental Venous Anomaly, a common, benign vascular malformation that is often present with a cerebral cavernous malformation.
	
	\item[SWS] Sturge-Weber Syndrome, a sporadic vascular malformation disorder characterized by a mosaic port-wine stain on the face and neurological involvement. 
	
	\item[PWS] Port-Wine Stain, a benign vascular malformation typically on the face. May form independently, or as part of Sturge-Weber syndrome. 
	
	\item[FFPE] Formalin-Fixed Paraffin-Embedded, a common tissue fixation strategy in pathology. The fixation process partially degrades DNA and RNA in the tissue. 
	
	\item[COSMIC] Catalog Of Somatic Mutations In Cancer, a web database containing somatic mutations that have been identified in cancers. 
	
	\item[LOH] Loss Of Heterozygosity, the loss of one allele at a heterozygous site. LOH can result from various genetic mechanisms (e.g. deletion, recombination, mitotic nondisjunction).
	
\end{symbollist}
