\acknowledgements
First, I would like to acknowledge everyone that donated tissue and participated in these studies. This is ultimately for them, and without them, would not be possible. 
\\ \\
I would like to thank my advisor, Douglas Marchuk, for his mentorship and guidance throughout my research and for his boundless enthusiasm for science. I am also thankful for the support of the current and former members of the Marchuk Lab: Han Kyu Lee, Sarah Wetzel-Strong, Leandro Barbosa Do Prado, Matthew Detter, Francesca Galeffi, Carol Gallione, Christian Benavides, Mary Patrucco, Catherine Neilson, and Erin Griffin. I would also like to thank Craig Lowe for his mentorship and for always welcoming me into his lab. I also thank the members of the Lowe Lab for always being open to talk science and for their friendship: Eric Au, Riley Mangan, Christiana Fauci, Luke Bartelt, Chelsea Shoben, Seth Weaver, Yanting Luo, Kat Bendt, and Juliana Carvalho. I would also like to acknowledge the members of my thesis advisory committee for their advice throughout these studies: Beth Sullivan, Michael Hauser, Timothy Reddy, and Craig Lowe. 

I could not have completed any of this work without the support of our many collaborators. In particular I would like to acknowledge the contributions of Aileen Ren, Mark Kahn, Issam Awad, Mark Ginsberg, Paul Oh, Michael Lawton, Marie Faughnan, Romuald Girard, Amy Akers, the Angioma Alliance Foundation, and the Cure HHT Foundation. I would also like to acknowledge funding from the National Heart Lung and Blood Institute and the American Society for Human Genetics.

I would like to acknowledge multiple core research facilities at Duke University that were critical for the completion of this work: Sequencing and Genomic Technologies Core (Nicolas Devos, Holly Dressman, Laura-Leigh Rowlette, Fangfei Ye), Duke Human Vaccine Institute Flow Cytometry Core (Derek Cain, Evan Trudeau, Steven Slater), and the Duke Center for Genomic and Computational Biology High-throughput Applied Research Data Analysis Cluster (Hilmar Lapp, Gregory Steffen, Dan Somers). 
\\ \\
Lastly, I thank my family and friends for their loving support throughout this journey. 