\biography

Daniel Aaron Snellings was born in La Plata, Maryland on the 23\textsuperscript{rd} of October, 1995. He graduated from the Pennsylvania State University in 2017 with a Bachelor of Science in Biochemistry and Molecular Biology. Daniel entered Duke in 2017 through the Cell and Molecular Biology graduate program and joined the lab of Douglas A. Marchuk in 2018. Daniel was awarded a Ruth L. Kirschstein predoctoral fellowship from the NHLBI in 2020. In 2021, Daniel was selected as a featured plenary speaker for the annual meeting of the American Society of Human Genetics and there was selected as a semifinalist for the Charles J. Epstein award. Daniel received a Ph.D. in Molecular Genetics and Microbiology from Duke University in 2022. In April 2022, Daniel began a postdoctoral fellowship in the lab of Christopher A. Walsh at Boston Children's Hospital.
\\\\
{\Large \textbf{Publications}} \\
\textbf{D. A. Snellings}, et al. (2021) \italicize{bioR$\chi$iv}\\ % ADD CITATION WHEN ACCEPTED TO NAT CARDIO RESC
\textbf{D. A. Snellings}, et al. (2021) \italicize{Circ Res}\\ \nocite{snellings2021}
A. A. Ren \& \textbf{D. A. Snellings}, et al. (2021) \italicize{Nature}\\ \nocite{ren2021}
\textbf{D. A. Snellings}, et al. (2019) \italicize{Am J Hum Genet}\\ \nocite{snellings2019}
J. Koskimaki, et al. (2019) \italicize{Acta Neuropathol Commun}\\ \nocite{koskimaki2019}
M. R. Detter, \textbf{D. A. Snellings}, D. A. Marchuk. (2018) \italicize{Circ Res}\\ \nocite{detter2018}
C. Wang, et al. (2018) \italicize{PLoS One} \nocite{wang2018}